---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# amltargets

<!-- badges: start -->
![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange)
<!-- badges: end -->

`amltargets` extends `targets` with a single high-leverage factory,
`tar_aml_job()`, to offload expensive targets to Azure Machine Learning (Azure ML)
compute clusters.

The package keeps your normal `targets` workflow while adding:

- Asynchronous Azure ML command job submission
- Polling and status tracking from your local R session
- Resume after interruption using JSON run tokens in `_targets/azure_tokens/`
- Hybrid execution where only selected targets run in Azure ML

## Why amltargets?

If your pipeline has a few targets that are too heavy for a laptop but the rest is
fast locally, `amltargets` gives you a clean split:

- Keep lightweight orchestration local with `targets`
- Send expensive compute to Azure ML clusters
- Preserve static dependency analysis and normal `targets` ergonomics

## Installation

Install from GitHub with `pak` (recommended):

```{r install, eval=FALSE}
# install.packages("pak")
pak::pak("jeffery-leirness/amltargets")
```

## Prerequisites

Before running `tar_aml_job()`, make sure you have:

- Azure CLI installed and authenticated (`az login`)
- An Azure ML workspace and compute cluster
- A registered Azure ML environment containing R and your dependencies
- A datastore path for the shared `_targets` cache, for example:
  `azureml://datastores/workspaceblobstore/paths/pipeline/_targets/`

You can set workspace defaults once in `.Renviron`:

```ini
AZURE_RESOURCE_GROUP=my-resource-group
AZURE_ML_WORKSPACE=my-ml-workspace
```

### Mount `_targets` on your Compute Instance

To share the same cache between local orchestration and cluster jobs, mount the
blob datastore path locally and symlink it into your project as `_targets/`.

```bash
mkdir -p "$HOME/amltargets-mounts/pipeline-targets"

az mount \
  --path "azureml://datastores/workspaceblobstore/paths/pipeline/_targets/" \
  --mount-point "$HOME/amltargets-mounts/pipeline-targets" \
  --resource-group my-resource-group \
  --workspace-name my-ml-workspace

cd /path/to/your/project
ln -sfn "$HOME/amltargets-mounts/pipeline-targets" _targets
ls -ld _targets
```

If `az mount` is not recognized, update Azure CLI and the ML extension, then
rerun the command.

## Usage

Use `tar_aml_job()` anywhere you would normally use `targets::tar_target()` for
compute-heavy steps.

```{r usage, eval=FALSE}
library(targets)
library(amltargets)

DATASTORE <- "azureml://datastores/workspaceblobstore/paths/pipeline/_targets/"

list(
  tar_target(raw_data, mtcars),

  tar_target(
    splits,
    {
      set.seed(123)
      train_idx <- sample(seq_len(nrow(raw_data)), floor(0.8 * nrow(raw_data)))
      list(
        train = raw_data[train_idx, ],
        test = raw_data[-train_idx, ]
      )
    }
  ),

  tar_aml_job(
    model_fit,
    lm(mpg ~ wt + cyl, data = splits$train),
    cluster = "gpu-cluster",
    datastore_path = DATASTORE,
    environment = "azureml:r-tidyverse@latest",
    poll_interval = 60L
  ),

  tar_target(
    mse,
    {
      preds <- predict(model_fit, newdata = splits$test)
      mean((splits$test$mpg - preds)^2)
    }
  )
)
```

Run as usual:

```{r run, eval=FALSE}
tar_make()
```

## Execution modes

At runtime, the orchestrator automatically selects one mode:

- Hybrid (default): submit to Azure ML and poll until terminal state
- Cluster mode: if `AZUREML_RUN_ID` is set, evaluate directly on-cluster
- Forced local mode: if `TAR_AZURE_ML_LOCAL=true` or
  `options(tar_azure_ml_local = TRUE)`

For rapid local tests without cloud submission:

```{r local-mode, eval=FALSE}
Sys.setenv(TAR_AZURE_ML_LOCAL = "true")
tar_make()
Sys.unsetenv("TAR_AZURE_ML_LOCAL")
```

## How resilience works

When an Azure ML job is submitted, `amltargets` writes a token file containing
the run ID. If `tar_make()` is interrupted (sleep, reboot, session exit), the
next run resumes tracking the existing Azure ML job instead of creating a
duplicate submission.

## Learn more

- Package vignette: `vignette("amltargets-run", package = "amltargets")`
- `targets` documentation: <https://docs.ropensci.org/targets/>

## License

MIT, see `LICENSE`.
