---
title: "Using amltargets::tar_aml_job(): Offloading Targets to Azure ML"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using amltargets::tar_aml_job(): Offloading Targets to Azure ML}
  %\VignetteDepends{amltargets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

`amltargets` extends the [`targets`](https://docs.ropensci.org/targets/) package
with a single function, `tar_aml_job()`, that offloads a target's computation
to an **Azure ML Compute Cluster** as an asynchronous Command Job.
The local R session submits the job, then polls until it finishes —
and if your laptop sleeps or reboots in the meantime, the pipeline picks up
exactly where it left off.

## Prerequisites

Before using this package you need:

1. **Azure CLI** installed and authenticated (`az login`).
2. An **Azure ML workspace** with at least one compute cluster provisioned.
3. An **Azure ML environment** that has R (and your package dependencies)
   installed. See [Registering an R environment](#sec-environment) below.
4. A **split-storage setup** with two Azure storage resources:
   - **`workspacefileshare`** (Azure Files) — mounted on both the local Compute
     Instance and the cluster. Your R scripts and `_targets.R` live here.
     The job YAML sets `code` to the current working directory (`getwd()`),
     which Azure ML snapshots as the job source.
   - **`workspaceblobstore`** (Azure Blob Storage) — used exclusively for the
     `_targets/` pipeline cache. It is attached to the cluster container as a
     read-write output mount, and the job symlinks it to `_targets/` in the
     working directory before R starts. Pass its datastore URI as the
     `datastore_path` argument.
   - The `datastore_path` URI takes the form
     `"azureml://datastores/workspaceblobstore/paths/<your/path>/"`.
     Create the path under your datastore before running the pipeline.

## Mounting `_targets` on the local Compute Instance

The offloaded job already mounts the blob-backed datastore on the cluster side.
You should also mount the same datastore path on your local Azure ML Compute
Instance and symlink it into your project as `_targets/`.

This keeps `targets::tar_make()` and `targets::tar_read()` on the compute
instance pointed at the same cache location used by cluster jobs.

1. Mount your blob datastore path with `az ml datastore mount`:

```bash
az ml datastore mount \
  --path "azureml://datastores/workspaceblobstore/paths/pipeline/_targets/" \
  --mode rw_mount \
  --mount-point "$HOME/amltargets-mounts/pipeline-targets" \
  --resource-group my-resource-group \
  --workspace-name my-ml-workspace
```

2. From your project root (on `workspacefileshare`), symlink that mount as
   `_targets/`:

```bash
cd /path/to/your/project
ln -sfn "$HOME/amltargets-mounts/pipeline-targets" _targets
```

3. Verify the link:

```bash
ls -ld _targets
```

If `az ml datastore mount` is not recognized in your environment, update Azure CLI and
the ML extension, then rerun the mount command.

## Installation

```{r install}
# install.packages("pak")
pak::pkg_install("jeffery-leirness/amltargets")
```

## Setting up credentials

`tar_aml_job()` needs to know your Azure resource group and workspace name.
Rather than repeating them in every call, set them once in your project's
`.Renviron` file:

```ini
AZURE_RESOURCE_GROUP=my-resource-group
AZURE_ML_WORKSPACE=my-ml-workspace
```

Restart R (or call `readRenviron(".Renviron")`) to pick them up.
You can still override either value per-call if a target needs a different
workspace.

## Registering an R environment {#sec-environment}

Azure ML runs the job inside a Docker-based environment. You need to register
one that has R and your project's packages. The minimal approach is a curated
or custom environment defined in a YAML file.

Create `azureml/r-env.yaml` in your project:

```yaml
$schema: https://azuremlschemas.azureedge.net/latest/environment.schema.json
name: r-tidyverse
description: "R 4.5 + tidyverse for pipeline jobs"
image: rocker/tidyverse:4.5
```

Register it once:

```bash
az ml environment create \
  --file azureml/r-env.yaml \
  --resource-group my-resource-group \
  --workspace-name my-ml-workspace
```

Use `environment = "azureml:r-tidyverse@latest"` in your `tar_aml_job()`
calls.

## A full `_targets.R` example

Below is a complete pipeline that:

1. Reads data locally (a lightweight step, kept local).
2. Trains a model on a GPU cluster (expensive, offloaded to Azure ML).
3. Evaluates the model locally using the result written by the cluster.

```{r targets-file}
# _targets.R
library(targets)
library(amltargets)

# ── Shared Azure ML settings ─────────────────────────────────────────────────
# Set AZURE_RESOURCE_GROUP and AZURE_ML_WORKSPACE in .Renviron so they don't
# appear in every call below.
DATASTORE <- "azureml://datastores/workspaceblobstore/paths/pipeline/_targets/"

# ── Pipeline definition ───────────────────────────────────────────────────────
list(
  # Step 1 — load data locally
  tar_target(raw_data, command = mtcars),

  # Step 2 — split locally using base R
  tar_target(
    splits,
    command = {
      set.seed(123)
      # Create an 80/20 train/test split
      train_idx <- sample(seq_len(nrow(raw_data)), floor(0.8 * nrow(raw_data)))
      list(
        train = raw_data[train_idx, ],
        test = raw_data[-train_idx, ]
      )
    }
  ),

  # Step 3 — train on Azure ML GPU cluster  ← offloaded
  # The command expression references `splits`, so targets knows this target
  # depends on Step 2 and will wait for it before submitting the job.
  tar_aml_job(
    model_fit,
    command = {
      # Train a simple linear model predicting miles per gallon (mpg)
      # using weight (wt) and cylinders (cyl)
      lm(mpg ~ wt + cyl, data = splits$train)
    },
    cluster = "gpu-cluster",
    datastore_path = DATASTORE,
    environment = "azureml:r-tidyverse@latest",
    poll_interval = 60L
  ),

  # Step 4 — evaluate locally using base R predict()
  tar_target(
    metrics,
    command = {
      # Generate predictions on the test set
      preds <- predict(model_fit, newdata = splits$test)

      # Calculate Mean Squared Error (MSE)
      mse <- mean((splits$test$mpg - preds)^2)
      c(MSE = mse)
    }
  )
)
```

Run the pipeline as usual:

```{r run}
targets::tar_make()
```

`targets` handles dependency ordering automatically. The `model_fit` target
will not be submitted until `splits` is up to date. While the Azure ML job is
running, your console will display polling messages:

```
● run target model_fit
Submitted Azure ML job for target 'model_fit'. Run ID: dreamy_spoon_abc123xyz
[2026-07-01 14:03:00] Job dreamy_spoon_abc123xyz status: Preparing. Polling again in 60s...
[2026-07-01 14:04:00] Job dreamy_spoon_abc123xyz status: Running. Polling again in 60s...
[2026-07-01 14:09:00] Job dreamy_spoon_abc123xyz status: Completed.
● built target model_fit
```

## How reboot resilience works

When `tar_aml_job()` submits a job it writes a small JSON **token file**
to `<store>/azure_tokens/<target_name>.json` containing the Run ID:

```json
{ "run_id": "dreamy_spoon_abc123xyz" }
```

The store location is resolved at runtime by `targets::tar_path_store()`, so
it respects any custom store configured via `tar_config_set(store = ...)` or
`_targets.yaml`. By default this is the `_targets/` directory, which on your
local Compute Instance resolves through the `workspacefileshare` mount.

If your laptop sleeps, reboots, or `tar_make()` is interrupted, the token file
persists on the shared mount. On the next call to `tar_make()`, the orchestrator
finds the token and resumes polling the *existing* job — no duplicate submission.

Once the job reaches a terminal state (Completed, Failed, or Canceled), the
token is deleted automatically.

## Execution modes

The orchestrator chooses one of three modes at runtime:

| Mode | Trigger | Behaviour |
|---|---|---|
| **Hybrid** (default) | Neither flag is set | Submits to Azure ML, polls from local R session |
| **Cluster** | `AZUREML_RUN_ID` env var is set | Evaluates the command directly; used automatically inside the Azure ML job |
| **Forced local** | `TAR_AZURE_ML_LOCAL=true` or `options(tar_azure_ml_local = TRUE)` | Evaluates locally; useful for rapid local testing without submitting jobs |

To test your pipeline end-to-end locally before spending cloud compute:

```{r local-mode}
Sys.setenv(TAR_AZURE_ML_LOCAL = "true")
targets::tar_make()
Sys.unsetenv("TAR_AZURE_ML_LOCAL")
```

## Cluster-side job setup

When Azure ML picks up the Command Job, the actual entrypoint is a two-step
shell command generated by `write_job_yaml()`:

```bash
ln -s ${{outputs.cluster_targets_dir}} _targets && Rscript _azureml_<target_name>.R
```

The `ln -s` step symlinks the blob storage output mount (injected by Azure ML
as `${{outputs.cluster_targets_dir}}`) to `_targets/` in the cluster
container's working directory. This means `targets::tar_read()` and
`targets::tar_make()` on the cluster see the same cache as your local session.
The `code` field in the job YAML points to `getwd()`, so Azure ML snapshots
the current `workspacefileshare` project directory as the job source, and your scripts are
available on the cluster without any manual upload step.

Because that code snapshot now contains a symlink named `_targets`, add an
`.amlignore` file at the root of your project directory (next to `_targets.R`)
with this single line:

```text
_targets/
```

Azure ML respects `.amlignore` when packaging the code snapshot. Excluding
`_targets/` prevents Azure ML from traversing the symlink and attempting to zip
up the entire pipeline cache on every job submission, which keeps submission
latency low even when the blob-backed cache grows large. This lets you keep
your preferred workspace layout on `workspacefileshare` while using
`workspaceblobstore` as the high-capacity backend for `targets` data.

The generated `_azureml_<target_name>.R` worker script does three important
things before returning a value:

1. Sources `_targets.R` to load project functions, options, and package setup.
2. Uses `targets::tar_network()` and `targets::tar_load()` to load upstream
  dependencies required by the offloaded target.
3. Evaluates your command expression, then writes the result to
  `_targets/azure_transfers/<target_name>.rds` so the local orchestrator can
  read it and apply the requested target format.

For this to work, the cluster node must also be able to:

- **Load all packages** referenced in `command`. Include them in your Azure ML
  environment image or install them via a job `setup_script`.
- **Evaluate target code from your project**. Because `_targets.R` is sourced
  automatically by the worker script, your command usually does not need an
  explicit `targets::tar_source()` call.

A command block that uses upstream dependencies and project helpers directly:

```{r self-contained}
DATASTORE <- "azureml://datastores/workspaceblobstore/paths/pipeline/_targets/"

tar_aml_job(
  model_fit,
  command = lm(mpg ~ wt + cyl, data = splits$train),
  cluster = "gpu-cluster",
  datastore_path = DATASTORE,
  environment = "azureml:r-tidyverse@latest"
)
```

## Passing `tar_target()` options

`tar_aml_job()` forwards `...` to `targets::tar_target_raw()`, so any
standard target option works:

```{r tar-options}
tar_aml_job(
  big_embedding,
  compute_embeddings(corpus),
  cluster = "gpu-cluster",
  datastore_path = DATASTORE,
  environment = "azureml:r-tidymodels:1",
  format = "qs",   # fast serialisation for large objects
  priority = 1       # run this target first
)
```

## Inspecting the pipeline manifest

Because `tar_aml_job()` uses non-standard evaluation, `targets` can still
perform static code analysis and build the full dependency graph:

```{r manifest}
targets::tar_manifest()
targets::tar_visnetwork()
```

The `model_fit` target will appear in the graph with `splits` as an upstream
dependency, even though `model_fit` is evaluated on Azure ML while `splits`
is built locally.
